---
title: "Tidyverse"
subtitle: "In-Class Activity"
author: "Dyer"
format: 
  html:
    echo: true 
    code-tools: true
abstract: This in-class activity will be used in the lecture to facilitate the use of several aspects relevant to applying the 'verbs' of data manipulation using the `tidyverse` library.
---

## The Question

> Load in the provided data set to answer the following questions. In the space below create a table of the following data and in-text references to the table.
>
> 1.  Make a table showing minimum, mean, and maximum water turbidity for each month in the dataset, and\
> 2.  Describe the daytime air temperatures at the Rice Rivers Center for February, 2014 by day of the week

## The Data

For this topic we will use some example data from the [Rice Rivers Center](https://ricerivers.vcu.edu). These data represent both atmospheric and water data collected from instrumentation on-site. I have stored these data in a spreadsheet that is shared on Google Drive as a CSV file.

```{r}
library( tidyverse )
library( lubridate )
library( knitr )
library( kableExtra )

url <- "https://docs.google.com/spreadsheets/d/1Mk1YGH9LqjF7drJE-td1G_JkdADOU0eMlrP01WFBT8s/pub?gid=0&single=true&output=csv"
rice <- read_csv( url )
summary( rice )
```


```{r}
rice |> 
  select( AirTempF, RecordID, everything()  )
```




These data represent measurements taken every 15 minutes, 24 hours a day, 7 days a week, 365 days a year. If you look at the summary of the data above, you will see several things, including:

-   Date and time objects are `character`
-   Some measurements are in Standard and some in Imperial with units in the same file include both °F and °C, as well as measurements in meters, feet, and inches. In fact, there are duplication of data columns in different units (guess what kind of correlation they might have...)

```{r}
rice |>
  select( DateTime, AirTempF, Turbidity_ntu ) |>
  mutate( Date = mdy_hms( DateTime ) ) |> 
  select( -DateTime ) |> 
  mutate( Month = month( Date ) ) |> 
  mutate( Weekday = wday( Date, 
                          label = TRUE,
                          abbr = FALSE ) ) |> 
  mutate( Temp = (AirTempF - 32) * 5/9 ) |> 
  select( Month, 
          Weekday,
          Turbidity = Turbidity_ntu,
          `Air Temperature (°C)` = Temp ) -> data

```


### Examples 

HEre are some examples 



```{r}
#| label: tbl-turbidity
#| echo: false
#| warning: false
#| tbl-cap: Minimum, mean, and maximum turbidity sampled from the pier at the Rice Rivers Center in 202alkdasfj;lakdjfg;alskfj

# this is a comment

data |>
  filter( `Air Temperature (°C)` > 0 ) |>
  filter( !is.na( Turbidity ) ) |> 
  filter( Month == 2 ) |> 
  group_by( Weekday ) |> 
  summarize( Minimum = min( Turbidity ), 
             Average = mean( Turbidity ),
             Maximum = max( Turbidity ) ) |> 
  kable() 


```



There are several cool things to see in what ialksdajf; lkjasdf;kj a;sldfj ;asldkfj;asdlfk ja; ; see @tbl-turbidity.








## The Results

The data from the Rice Rivers Center was evaluated for water turbidity and daily air temperature.

### Turbidity

```{r}

data |>
  group_by( Month ) |>
  summarize( Minimum = min( Turbidity, na.rm=TRUE ),
             Mean = mean( Turbidity, na.rm=TRUE), 
             Maximum = max( Turbidity, na.rm=TRUE ) ) |>
  kable( digits = 1 ) |>
  kable_paper() 


```

### Daily Air Temperature

```{r}
# Get day-of-the-week air temperature



```






```{r}
library( tidyr )

data |>
  select( -Month, -Weekday ) |> 
  pivot_longer( 
    cols = c(Turbidity, `Air Temperature (°C)`) 
    ) |> 
  group_by( name ) |> 
  summarize( Minimum = min( value, na.rm=TRUE ),
             Mean = mean( value, na.rm=TRUE ), 
             Maximum = max( value, na.rm=TRUE ) )

```
